---
title: "NARS data download"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(readr)
library(dplyr)
library(purrr)
```

Read in the table created with `get_nars_links.R`. 

```{r}
nars_data <- read_csv("data/nars_data_table.csv")
```

List all the values in the indicator column

```{r}
unique(nars_data$Indicator)
```

Get the subset of files that are associated with Water Chemistry data

```{r}
wq_data <- filter(nars_data, Indicator == "Water Chemistry")
```

```{r}
wq_data$data_link
```
Download all the csv files in the data links column using basename as object name

This would make one data frame if they all have the same columns....
```{r}
# basename(wq_data$data_link, )
# woops theres one that is xlsx
wq_df <- wq_data$data_link[1:2] %>%
  map_df(read_csv, .id = "filename")
```

This will make a list where the input from each file is separate, and then can be iterated over. 5th link is xlsx. not sure what is going on with 3rd file

```{r}
wq_dfs_list <- wq_data$data_link[c(1:2, 4, 6:13)] %>%
  map(read_csv)
```

