---
title: "NARS data download"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(readr)
library(dplyr)
library(purrr)
library(fs)
```

Read in the csv table created with `get_nars_links.R`. 

```{r}
nars_data <- read_csv("data/nars_data_table.csv")
nars_data$filename <- basename(nars_data$data_link)
```

Download all the csv files in the data links column into a folder called data (needs to exist first)

```{r}
map2(.x = nars_data$data_link,
     .y = file.path("data", basename(nars_data$data_link)), 
     ~download.file(.x, .y))
```

List all the values in the indicator column

```{r}
unique(nars_data$Indicator)
```


Get the subset of files that are associated with Water Chemistry data

```{r}
wq_data <- filter(nars_data, Indicator == "Water Chemistry")
```

This would make one data frame if they all have the same columns....

```{r}
# basename(wq_data$data_link, )
# woops theres one that is xlsx
wq_df <- wq_data$data_link[1:2] %>%
  map_df(read_csv, .id = "filename")
```

This will make a list where the input from each file is separate, and then can be iterated over. 5th link is xlsx. not sure what is going on with 3rd file

```{r}
wq_dfs_list <- wq_data$data_link[c(1:2, 4, 6:13)] %>%
  map(read_csv)
```

Subset to just Site information tables

```{r}
site_info_data <- nars_data %>% 
  filter(Indicator == "Site Information")
```

2nd file is weird - read in 1st and 3rd files

```{r}
site_dfs_list <- site_info_data$data_link[c(1,3)] %>%
  map(read_csv)
```

## Map

Mapping data from first table

```{r}
library(leaflet)
library(sf)
site_dfs_list[[1]] %>%
  st_as_sf(coords = c("LON_DD", "LAT_DD"), crs = 4269) %>%
  leaflet() %>%
  addTiles(group = "Open Street Maps") %>%
  addProviderTiles(providers$Esri.WorldImagery, group = "Esri Imagery") %>%
  addMarkers(popup = ~LAKENAME, clusterOptions = markerClusterOptions(),
             group = "sites") %>%
  addLayersControl(overlayGroups = c("sites"), 
                   baseGroups = c("Open Street Maps", "Esri Imagery"))
  
```

