---
title: "Get NLCD data"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

FedData package has a function for downloading NLCD tiles. Let's see if it works.

```{r, message=FALSE}
# install.packages("FedData")
library(FedData)
library(sf)
library(leaflet)
library(readr)
library(dplyr)
```

Using wetlands site info file, read in from internet

```{r, message=FALSE}
w2011_siteinfo <- read_csv("https://www.epa.gov/sites/production/files/2016-10/nwca2011_siteinfo.csv")
# # or using code below
# nars_data <- read_csv("data/nars_data_table.csv")
# nars_data$filename <- basename(nars_data$data_link)
# 
# w2011_siteinfo <- nars_data %>% 
#   filter(Survey == "Wetlands 2011",
#          Indicator == "Site Information") %>%
#   pull(filename) %>% 
#   tail(1) %>% # aachar and siteinfo files
#   paste0("data/", .) %>%
#   read_csv()
```

make into a spatial object assuming these columns are the correct coordinates for sites and in WSG 84. 

```{r}
template <- w2011_siteinfo %>%
  st_as_sf(coords = c("AA_CENTER_LON", "AA_CENTER_LAT"), crs = 4326)
```

Check out where the first few are
```{r}
template[1:4,] %>%
  st_transform(crs = 4326) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```

Plot all site locations

```{r}
plot(template$geometry)
```

Make buffer around first point
```{r}
pt1_buffer <- template %>%
  dplyr::slice(1) %>% 
  st_buffer(0.01) # units decimal degrees. i think this is approx 1000m
```

Plot buffer
```{r}
leaflet(pt1_buffer) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addPolygons()
```

Use buffer in get nlcd function. 
```{r}
nlcd_nars <- get_nlcd(as(pt1_buffer, "Spatial"), # wants a spatial sp object not sf 
                      label = "first wetland site", 
                      dataset = "landcover",
                      force.redo = TRUE)
```

```{r}
nlcd_nars %>% 
  plot()
```

```{r}
raster::values(nlcd_nars) %>% unique()
# nlcd_landcover_pam this is a terrible format
```

Unique values 90, 42, 21, 95, 71
https://www.mrlc.gov/nlcd11_leg.php

```{r}
# lc_legend_df <- read.table("nlcd_legend.txt", stringsAsFactors = FALSE, sep = ",")
# lc_legend_df<- subset(lc_legend_df,COUNT>0) #subset the data to remove unused rows (but this is from 2006)
```

```{r}
lc_legend_df <- readr::read_csv("nlcd_legend_2011.csv")
```

```{r}
leaflet(pt1_buffer) %>%
  addProviderTiles(providers$Esri.WorldImagery) %>%
  addPolygons() %>%
  addRasterImage(nlcd_nars)
```

Mask raster using buffer object after transforming to same CRS. 

```{r}
pt1_buffer_prj <- st_transform(pt1_buffer, crs = proj4string(nlcd_nars))
nlcd_nars_mask <- raster::mask(nlcd_nars, as(pt1_buffer_prj, "Spatial"))
```

```{r}
plot(nlcd_nars_mask)
```

tabulate number of cells in each type

```{r}
raster::freq(nlcd_nars_mask)
```
 Merge with legend to see land cover types
```{r}
raster::freq(nlcd_nars_mask) %>%
  as.data.frame() %>%
  dplyr::filter(!is.na(value)) %>%
 dplyr::left_join(lc_legend_df, by = c("value" = "Class"))
```

