---
title: "Streams 2008-20099 NARS data"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and list of NARS survey data files
```{r}
# install.packages('pacman')
pacman::p_load(tidyverse)

nars_data <- read_csv("data/nars_data_table.csv") %>%
  mutate(filename = basename(.$data_link))
```

Print unique NARS survey names
```{r}
nars_data %>% pull(Survey) %>% unique()
```

List data types within Rivers and Streams 2008-2009
```{r}
nars_data %>% 
  filter(Survey == "Rivers and Streams 2008-2009") %>%
  pull(Indicator) %>% unique()
```

List file names within Water chemistry
```{r}
nars_data %>% 
  filter(Survey == "Rivers and Streams 2008-2009") %>%
  pull(Data) %>% unique()
```

Water chemistry data (DOC, TN, TP). 2 files for chemistry - chem and chemcond
Raw water chem data is in "chem" file; "chemcond" is Indicator Values and Condition Class
```{r}
streams09 <- nars_data %>%
  filter(Survey %in% c("Rivers and Streams 2008-2009"), 
         Indicator %in% c("Water Chemistry"),
         str_detect(Data, 'Water Chemistry - Data')) %>% # Select for the raw data file
  pull(filename) %>% paste0("data/", .) %>%
  read_csv()
```

List variable names in water quality file
```{r}
names(streams09)
```

Select focal variables (site information, DOC, TN, TP, and associated data flags)
```{r}
streams09_CNP <- streams09 %>% 
  select(YEAR, UID, SITE_ID, DOC, DOC_ALERT, NTL, NTL_ALERT, PTL, PTL_ALERT)
```

Quick check of frequency of constituent flags
```{r}
streams09_CNP %>% 
  mutate(DOC_Flag = ifelse(!is.na(DOC_ALERT), "Yes", "No"),
         TN_Flag = ifelse(!is.na(NTL_ALERT), "Yes", "No"), 
         TP_Flag = ifelse(!is.na(PTL_ALERT), "Yes", "No")) %>%
  group_by(DOC_Flag, TN_Flag, TP_Flag) %>% count() 
```

Calculate CN, CP, NP (no QC yet)
```{r}
streams09_CNP <- streams09_CNP %>%
  mutate(CN = (DOC / (NTL/1000)) * (14.01/12.01), # Convert NTL to mg/L
         CP = (DOC / (PTL/1000)) * (30.97/12.01), # Convert PTL to mg/L
         NP = ((NTL/1000) / (PTL/1000)) * (30.97/14.01))  # Convert NTL, PTL to mg/L
```

Quick histograms of all nutrient constituents (DOC, TN, TP) and ratios
```{r}
streamCNP_long <- streams09_CNP %>%
  select(-DOC_ALERT, -NTL_ALERT, -PTL_ALERT) %>% 
  gather(Variable, Value, DOC:NP) 

# DOC, TN, TP
ggplot(subset(streamCNP_long, Variable %in% c('DOC', 'NTL', 'PTL')), 
       aes(x = Value, fill= Variable)) + 
  geom_histogram(bins = 100) + guides(fill = FALSE) +
  facet_wrap(. ~ Variable, scales='free')

# Molar CN, CP, NP
ggplot(subset(streamCNP_long, Variable %in% c('CN', 'CP', 'NP')), 
       aes(x = Value, fill= Variable)) + 
  geom_histogram(bins = 100) + guides(fill = FALSE) +
  facet_wrap(. ~ Variable, scales='free')
```

Scatterplots of TN, TP data 
```{r}
# Full extent of data
ggplot(streams09_CNP, aes(y = PTL, x = NTL)) + 
  geom_point() + 
  labs(y = (expression(Total~Phosphorus~(mu*g~L^-1))),
       x = (expression(Total~Nitrogen~(mu*g~L^-1))))

# Truncated axes
ggplot(streams09_CNP, aes(y = PTL, x = NTL)) + 
  geom_point() + 
  labs(y = (expression(Total~Phosphorus~(mu*g~L^-1))),
       x = (expression(Total~Nitrogen~(mu*g~L^-1)))) +
  scale_y_continuous(limits=c(0,2000)) +
  scale_x_continuous(limits=c(0,10000))
```

##Landscape data
```{r}
# check 2 parsing failures
streams09_landscape <- nars_data %>% 
  filter(Survey == "Rivers and Streams 2008-2009",
         Indicator == "Landscape Data",
         str_detect(Data, ' Landscape Metrics')) %>% # Other file is shapefiles
  pull(filename) %>% paste0("data/", .) %>%
  read_csv()

names(streams09_landscape)

# Select focal landscape variables (temperatures, land cover)
streams09_landscape <- streams09_landscape %>%
  select(UID, SITE_ID, TMAX_ANN:TMIN_JUL, PCT_AG:PCT_SHRUB_GRASS, NHDWAT_PCT_IMPERV)
```

##Site data: includes Ecoregion, EPA region, lat/long
```{r}
streams09_siteinfo <- nars_data %>% 
  filter(Survey == "Rivers and Streams 2008-2009",
         Indicator == "Site Information",
         str_detect(Data, 'NRSA 0809 Site Information')) %>%
  pull(filename) %>% 
  paste0("data/", .) %>%
  read_csv()

# "External site data" includes 11 more variables
streams09_Exsiteinfo <- nars_data %>% 
  filter(Survey == "Rivers and Streams 2008-2009",
         Indicator == "Site Information",
         str_detect(Data, 'External Data')) %>%
  pull(filename) %>% 
  paste0("data/", .) %>%
  read_csv()
```
